### 2019-03-03 report/final-report.tex
added symmetric synthesis to methods section

Figure~\Cref{fig:schematic-diagram-synthesis} shows a high-level,
schematic diagram for \Optician{}.
First, \Optician{} uses the function \ToDNFRegex{} to convert the input
regular expressions into DNF regular expressions.  Next,
\SynthDNFLens{} performs type-directed synthesis on these DNF regular
expressions and the input examples to synthesize a DNF lens.  Finally, this DNF lens is
converted back into a regular lens with the function \ToLens{}, and returned to the user.

@@

>>>>

\Cref{fig:schematic-diagram-synthesis} shows a high-level,
schematic diagram for \Optician{}.
First, \Optician{} uses the function \ToDNFRegex{} to convert the input
regular expressions into DNF regular expressions.  Next,
\SynthDNFLens{} performs type-directed synthesis on these DNF regular
expressions and the input examples to synthesize a DNF lens.  Finally, this DNF lens is
converted back into a regular lens with the function \ToLens{}, and returned to the user.

@@

\subsection{Simple symmetric lens synthesis}
Although it is very useful to be able to synthesize bijective or
quotient lenses between data formats, such lenses do not fully solve the
problem. 
Bijective and quotient lenses require the two data formats to have
either precisely or morally the same information content. 
Many related data formats in practice have overlapping
information: some information may be present in one format but not the
other and vice versa.

Symmetric lenses~\cite{HofmannPierceWagner10:POPL} address this
limitation.  General symmetric lenses introduce the notion of 
\textit{complement} that stores the information necessary to fully
reconstruct one format from the other.  Because such complements would
complicate the synthesis specificaiton problem, we instead define a
more restricted language of \textit{simple symmetric lenses}, which
are the largest class of symmetric lenses that do not rely on
persistent internal state to synchronise between two related data
formats, a property called \textit{forgetfulness}.  (Instead of a
complement, simple symmetric lenses rely on defaults to replace
missing information.)

A challenge in adopting the type-directed synthesis approach to simple
symmetric lenses is the number of such lenses.  Whereas the number of
bijective lenses between two formats is typically tiny, the
number of simple symmetric lenses is typically enormous. 
If a na\"ive  search algorithm just selects the first simple symmetric
lens it finds, the returned lens will generally not be the one the
user wanted. Symmetric synthesis requires a new principle for
identifying “more likely” lenses and a more sophisticated synthesis
algorithm that uses this principle to search the space more
intelligently.

For these, we turn to information theory. We consider ``likely'' lenses to
be ones that propagate ``a lot'' of information 
from the left data format to the right and vice versa. Conversely, ``unlikely''
lenses are ones that require a large amount of additional information to recover
one of the formats given the other. By default, our synthesis algorithm prefers
lenses that propagate more information.
This preference is formalized using \emph{stochastic regular expressions}
(SREs)~\cite{stoch-rnn,stoch-def}, which simultaneously define a set of
strings and a probability distribution over those strings.  Using this
probability distribution, we can calculate the likelihood of a given
lens.
We also allow users to override the default mechanism for
calculating the information content of a SRE by
asserting that certain strings are \emph{essential} or \emph{irrelevant},
forcing certain data to either be retained or discarded during the
transformations.

\begin{figure}
 \centering
  \includegraphics[width=.63\textwidth]{high-level-algorithm.pdf}
  \vspace{-2ex}
  \caption{Schematic diagram for the simple symmetric lens synthesis algorithm.
    The user provides regular expressions \BRegex and \BRegexAlt and a set of
    examples \Examples{} as input. \Expand first converts \BRegex and \BRegexAlt
    to stochastic regular expressions \Regex and \RegexAlt with default
    probabilities. It then finds pairs of stochastic regular expressions
    equivalent to \Regex and \RegexAlt and iteratively proposes them to
    \GreedySynth. \GreedySynth finds a lens typed between the supplied SREs.
    When the algorithm finds a likely lens, it returns it.}
  \label{fig:high-level-algorithm}
\end{figure}

With simple symmetric lenses and this SRE-based likelihood measure in hand, we
propose a new algorithm for synthesizing likely lenses. At its core, the
algorithm performs a type-directed search between descriptions of the data
formats, measuring success using the likelihood measure.

Interesting complications arise from the need to deal with regular-expression
equivalences. There are infinitely many regular expressions equivalent to a
given one, and the lens returned by a type-directed search will in general
depend on which of the possible representations are chosen for its source and
target formats. Moreover, certain lenses may not be well-typed unless the
format representations are replaced by equivalent ones:
in general, the synthesis algorithm has to search through
equivalent regular expression types to find the most likely lens. To tame
this complexity, we divide the synthesis algorithm into two communicating search
procedures (Figure~\ref{fig:high-level-algorithm}), following~\cite{optician}.
The first, \Expand, uses rewriting rules to propose new pairs of
stochastic regular expressions equivalent to the original pair. The second,
\GreedySynth, uses a greedy, type-directed
algorithm to find a simple symmetric lens between input SRE pairs, returning the lens and
its likelihood score to \Expand. The whole synthesis algorithm heuristically
terminates when a sufficiently likely lens is found.
